Physics of Tidal Dissipation in Early-Type Stars and White Dwarfs: Hydrodynamical Simulations of Internal Gravity Wave Breaking in Stellar Envelopes

### 2020-02-19 writeups/Su_IGW_break.tex
small

\title[Physics of Tidal Dissipation]{Physics of Tidal Dissipation in Early-Type
Stars and White Dwarfs: Hydrodynamical Simulations of Internal Gravity Wave
Breaking in Stellar Envelopes}
\author[Y. Su et\ al.]{
Yubo Su,$^1$,
Daniel Lecoanet,$^2$
Dong Lai$^{1,3}$
\\
$^1$ Cornell Center for Astrophysics and Planetary Science, Department of
Astronomy, Cornell University, Ithaca, NY 14853, USA
\\
$^2$ Princeton Center for Theoretical Science, Princeton University, Princeton,
NJ 08544, USA
\\
$^3$ Department of Astrophysical Sciences, Princeton University, Princeton NJ
08544, USA
\\
$^4$ Tsung-Dao Lee Institute, Shanghai Jiao Tong University, Shanghai, 200240, China
}

@@

\begin{abstract}
    In binaries composed of early-type stars and white dwarfs, the dominant
    tidal process involves the excitation of internal gravity waves (IGWs),
    which propagate towards the stellar surface, and their dissipation via
    nonlinear wave breaking. We perform 2D hydrodynamical simulations of this
    wave breaking process in a stratified isothermal atmosphere. We find that,
    after an initial transient phase, the dissipation of the IGWs naturally
    generate a sharp critical layer, separating the lower stationary region
    (with no mean flow) and the upper ``synchronized'' region (with the mean
    flow velocity equal to the horizontal wave phase speed). While the critical
    layer is steepened by these waves, it is simultaneously broadened by
    Kelvin-Helmholtz instabilities such that, in steady state, the critical
    layer width is determined by the Richardson criterion. We study the
    absorption and reflection of incident waves off the critical layer and
    provide analytical formulae describing its evolution. The result of this
    study is important for characterizing the evolution of tidally heated white
    dwarfs and other binary stars.
\end{abstract}

>>>>

\title[Physics of Tidal Dissipation]{Physics of Tidal Dissipation in Early-Type
Stars and White Dwarfs: Hydrodynamical Simulations of Internal Gravity Wave
Breaking in Stellar Envelopes}
\author[Y. Su et\ al.]{
Yubo Su,$^1$,
Daniel Lecoanet,$^{2,3}$
Dong Lai$^{1,4}$
\\
$^1$ Cornell Center for Astrophysics and Planetary Science, Department of
Astronomy, Cornell University, Ithaca, NY 14853, USA
\\
$^2$ Princeton Center for Theoretical Science, Princeton University, Princeton,
NJ 08544, USA
\\
$^3$ Department of Astrophysical Sciences, Princeton University, Princeton NJ
08544, USA
\\
$^4$ Tsung-Dao Lee Institute, Shanghai Jiao Tong University, Shanghai, 200240, China
}

@@

\begin{abstract}
    In binaries composed of early-type stars and white dwarfs, the dominant
    tidal process involves the excitation of internal gravity waves (IGWs),
    which propagate towards the stellar surface, and their dissipation via
    nonlinear wave breaking. We perform 2D hydrodynamical simulations of this
    wave breaking process in a stratified, isothermal atmosphere. We find that,
    after an initial transient phase, the dissipation of the IGWs naturally
    generate a sharp critical layer, separating the lower stationary region
    (with no mean flow) and the upper ``synchronized'' region (with the mean
    flow velocity equal to the horizontal wave phase speed). While the critical
    layer is steepened by these waves, it is simultaneously broadened by
    Kelvin-Helmholtz instabilities such that, in steady state, the critical
    layer width is determined by the Richardson criterion. We study the
    absorption and reflection of incident waves off the critical layer and
    provide analytical formulae describing its evolution. The result of this
    study is important for characterizing the evolution of tidally heated white
    dwarfs and other binary stars.
\end{abstract}
